Non-word repetition in Yélî Dnye

### 2020-06-15 manuscript.Rmd
fixed little bugs, added freq ana and abstract

abstract: >
  ADD

@@

knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = FALSE, fig.pos = "T")
require("papaja")
library(kableExtra)
library(gdata)
library(ggplot2)

@@

## TODO

- double check demo: exact age missing: 78, 82, 83, 115, 119; rough age missing 78, there's a mixture of BGMF in sex, first born is missing, round age
- add frequency of sounds in stimuli in segments.xlsx from https://phoible.org/parameters
- then add that in stimuli to get minimum & average frequencies per stimulus
- then add analysis of NWR as a function of frequency
- draft discussion

@@

- look through for **mc**
- consent, "correct", "accuracy", "performance"
- help with table 1 (formatting)

@@

NWR has been used to seek answers to a variety of theoretical questions, including what  the links between phonology, working memory, and the lexicon are [@bowey2001nonword]. It is also frequently used as for applications, notably  as a diagnostic for language delays and disorders [@estes2007differences]. Since non-words can be generated in any language, it has attracted the attention of researchers working in multilingual and linguistically diverse environments,  particularly in Europe [@meir2016disentangling;@is08042009language].

@@

@@

```{r demo}
# read.csv("NWR-test-log_FINAL.csv",header=T,sep=";")->fulldemo
# fulldemo=fulldemo[,-2] #remove children's names
# fulldemo$dob=as.Date(as.character(fulldemo$DOB.ISO),format="%Y%m%d")
# fulldemo$test=as.Date(as.character(fulldemo$Date),format="%Y%m%d")
# fulldemo$age.exact=as.numeric(difftime(fulldemo$test,fulldemo$dob,unit="weeks"))/52.25
# fulldemo$Age=as.numeric(as.character(fulldemo$Age))
# plot(fulldemo$Age~fulldemo$age.exact)
# write.csv(fulldemo,"nwr-demo.csv")

@@

subset(demo,included=="yes" & Age <12 & !is.na(Age))->inc
#if age.exact is missing, use reported age
inc$age.exact[is.na(inc$age.exact)]<-inc$Age[is.na(inc$age.exact)]
#summary(inc)
table(inc$birthOrder)->mytab
```

A total `r dim(demo)[1]` children were tested, from `r length(levels(factor(demo$familyID)))` of families, in five hamlets. Some children could not be included for the following reasons: refused participation or failed to repeat items presented over headphones even after coaching (N=`r sum(demo$included=="didn\xd5t repeat")`), spoke too softly to allow offline coding (N=`r sum(demo$included=="spoke too softly")`). In addition, `r sum(demo$Age >12 | demo$included=="too old",na.rm=T)` teenagers were tested to put younger children at ease; their data is not included in analyses below. The remaining `r dim(inc)[1]` children (`r sum(inc$Sex=="G")` girls) were aged `r round(mean(inc$age.exact),2)` years (range `r round(min(inc$age.exact),2)`-`r round(max(inc$age.exact),2)` years). There were `r sum(inc$monolingual=="yes", na.rm=T)` children exposed only to Yélî in the home, `r sum(inc$monolingual=="no", na.rm=T)` children who were also exposed to another language in the home, and , `r sum(is.na(inc$monolingual))` for whom this information was missing. Maternal years of education averaged `r round(mean(inc$matEd,na.rm=T),2)` years (range `r round(min(inc$matEd,na.rm=T),2)`-`r round(max(inc$matEd,na.rm=T),2)` years; `r sum(is.na(inc$matEd))` children had this information missing).[^1] In terms of birth order, `r mytab[1]` were first borns, `r mytab[2]` second,  `r mytab[3]` third,  `r mytab[4]` forth,  `r mytab[5]` fifth, `r mytab[6]` sixth, and  `r mytab[7]` did not have this information.

@@

aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$age.exact,trnsc$attempt),mean)->pwcm_att
colnames(pwcm_att)<-c("ID","Age","Attempt","Word")

@@

aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.exact,trnsc$birthOrder,trnsc$matEd),mean)->overall
aggregate(trnsc$pc*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.exact,trnsc$birthOrder,trnsc$matEd),mean)->pc
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.exact,trnsc$birthOrder,trnsc$matEd),mean)->nld
overall=cbind(overall,pc$x,nld$x)
colnames(overall)<-c("ID","Sex","Age","BrthO","matEd","Word","Phoneme","NLD")

@@

```{r}

@@

We investigated patterns of deletion and substitution. Deletions were relatively rare, with only `r sum(del_tab[c(oralVs,nasalVs)])` vowels deleted, and `r sum(del_tab[c(simplexCs,complexCs)])` consonants.

@@

```{r bylength}
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$age.exact,trnsc$type),mean)->by_length
aggregate(trnsc$pc*100,by=list(trnsc$id,trnsc$age.exact,trnsc$type),mean)->pc
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$age.exact,trnsc$type),mean)->nld
by_length=cbind(by_length,pc$x,nld$x)
colnames(by_length)<-c("ID","Age","Length","Word","Phoneme","NLD")

@@

Our final exploratory analysis assessed whether variation in scores was structured by factors that vary across individuals. As shown in Figure 2, there was a greater deal of variance in earlier than later ages, with significantly higher NWR scores for older children  (Spearman's rank correlation, given inequality of variance, rho (`r corage$statistic`) = `r corage$estimate`, p = `r corage$p.value`). In contrast, there was no clear association between NWR scores, on the one hand, and sex (t (`r tsex$statistic`) = `r tsex$parameter`, p = `r tsex$p.value`), birth order  (`r corbo$statistic`) = `r corbo$estimate`, p = `r corbo$p.value`) or maternal education  (`r corme$statistic`) = `r corme$estimate`, p = `r corme$p.value`)

>>>>

abstract: >
   In nonword repetition (NWR) studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the local language. Accuracy is thought to reflect long-term phonological knowledge as well as online phonological working memory and flexible production patterns.In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory. This study contributes to three lines of research. First, we document that non-word items containing typologically rare sounds are repeated accurately less often that non-words containing more common sounds. Second, we document rather weak effects of item length, contributing to other research suggesting that length effects may be language-specific. Third, we do not find strong individual variation effects in this population, contrary to previous results documenting strong age-related effects. Together, these data provide a unique view of online phonological processing in a seldom-studied language, and contribute to both typological and language acquisition research.

@@

knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = FALSE, fig.pos = "T")
require("papaja")
library(kableExtra)
library(gdata)
library(ggplot2)
library(lme4)

@@

<!-- REMEMBER!!!! -->
<!-- phonetic characters: -->
<!-- - save excel as utf with save as -->
<!-- - do $ iconv -f utf-16le -t utf-8 stimuli.txt > stimconv.txt -->
<!-- - do $ iconv -f utf-16le -t utf-8 segments.txt > segconv.txt -->

@@

- look through this manuscript for bits of text to be written by **mc**
- help with table 1 (formatting & making it appear, because it's broken by the phonological representations)
- do you think Figure 2 should be done with phonetic symbols? If yes, could you try to implement that?
- double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for... If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
- double check my frequency entries in segments.xlsx using the search function in https://phoible.org/parameters; take a look at http://phoible.github.io/conventions/ in case you see something about double articulation (I didn't find tp or lBj)
- double check phonological implementation of stimuli in stimuli.xlsx 
- for both of the above, if you make any changes, they need to be re-saved as UTF16 txt file, and then converted (see remember above)
- double check also stim_seg_freq -- why are some items still with NA freq?
- consent, "correct", "accuracy", "performance"
- draft discussion

@@

NWR has been used to seek answers to a variety of theoretical questions, including what  the links between phonology, working memory, and the lexicon are [@bowey2001nonword], and to assess the extent to which apparent phonological constraints found in the lexicon affect online production [@gallagher2014identity]. It is also frequently used as for applications, notably  as a diagnostic for language delays and disorders [@estes2007differences]. Since non-words can be generated in any language, it has attracted the attention of researchers working in multilingual and linguistically diverse environments,  particularly in Europe [@meir2016disentangling;@is08042009language].

@@

Additionally, we looked up each phoneme in each target word in the Phoible database [@phoible]. We entered the number of languages in which each phoneme was found, as well as what percentage of languages in the database that represented. We could not find a small number of phonemes in the database, and treat those as NA values that do not contribute to the target-level global average phone cross-linguistic frequency, although an alternative would be to attribute them a very small number. * **mc** do you agree with this decision or shall we do the alternative?*

@@

```{r demo}
# read.csv("NWR-test-log_FINAL.csv",header=T,sep=";")->fulldemo
# fulldemo=fulldemo[,-2] #remove children's names
# fulldemo$dob=as.Date(as.character(fulldemo$DOB.ISO),format="%Y%m%d")
# fulldemo$test=as.Date(as.character(fulldemo$Date),format="%Y%m%d")
# fulldemo$age.rounded=as.numeric(difftime(fulldemo$test,fulldemo$dob,unit="weeks"))/52.25
# fulldemo$Age=as.numeric(as.character(fulldemo$Age))
# plot(fulldemo$Age~fulldemo$age.rounded)
# write.csv(fulldemo,"nwr-demo.csv")

@@

subset(demo,included=="yes" & Age <12 & !is.na(Age))->inc
#if age.rounded is missing, use reported age
inc$age.rounded[is.na(inc$age.rounded)]<-inc$Age[is.na(inc$age.rounded)]
#summary(inc)
table(inc$birthOrder)->mytab
```

A total `r dim(demo)[1]` children were tested, from `r length(levels(factor(demo$familyID)))` of families, in five hamlets. Some children could not be included for the following reasons: refused participation or failed to repeat items presented over headphones even after coaching (N=`r sum(demo$included=="didn\xd5t repeat")`), spoke too softly to allow offline coding (N=`r sum(demo$included=="spoke too softly")`). In addition, `r sum(demo$Age >12 | demo$included=="too old",na.rm=T)` teenagers were tested to put younger children at ease; their data is not included in analyses below. The remaining `r dim(inc)[1]` children (`r sum(inc$Sex=="G")` girls) were aged `r round(mean(inc$age.rounded),2)` years (range `r round(min(inc$age.rounded),2)`-`r round(max(inc$age.rounded),2)` years). There were `r sum(inc$monolingual=="yes", na.rm=T)` children exposed only to Yélî in the home, `r sum(inc$monolingual=="no", na.rm=T)` children who were also exposed to another language in the home, and , `r sum(is.na(inc$monolingual))` for whom this information was missing. Maternal years of education averaged `r round(mean(inc$matEd,na.rm=T),2)` years (range `r round(min(inc$matEd,na.rm=T),2)`-`r round(max(inc$matEd,na.rm=T),2)` years; `r sum(is.na(inc$matEd))` children had this information missing).[^1] In terms of birth order, `r mytab[1]` were first borns, `r mytab[2]` second,  `r mytab[3]` third,  `r mytab[4]` forth,  `r mytab[5]` fifth, `r mytab[6]` sixth, and  `r mytab[7]` did not have this information.

@@

aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$attempt),mean)->pwcm_att
colnames(pwcm_att)<-c("ID","Age","Attempt","Word")

@@

aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->overall
aggregate(trnsc$pc*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->pc
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$Sex,trnsc$age.rounded,trnsc$birthOrder,trnsc$matEd),mean)->nld
overall=cbind(overall,pc$x,nld$x)
colnames(overall)<-c("ID","Sex","Age","BrthO","matEd","Word","Phoneme","NLD")

@@

In this analysis, we were interested in variation in performance as a function of the average frequency with which sounds composing individual target words are found in languages over the world. To look at this, we fit a mixed logistic regression, in which the outcome variable was whether the non-word was correctly repeated or not. The fixed effects of interest was the average cross-linguistic phone frequency; we also included child age as a control fixed effect, and child ID & target ID as random variables.

```{r xling-freq-ana}
xl_fr=glmer(correct~avg_fr+age.rounded+(1|file) + (1|target),data=trnsc,family="binomial")
mycoeff=round(summary(xl_fr)$coefficients,3)
```

We could include 885 observations (**why -- shouldn't it be 1k??**), from 41 children producing in any given trial one of 41 potential target words. The analysis revealed a significant estimate for the average cross-linguistic frequency of phones in the target words (ß = `r mycoeff["avg_fr","Estimate"]`, SE ß = `r mycoeff["avg_fr","Std. Error"]`, p = `r mycoeff["avg_fr","Pr(>|z|)"]`): Target words with phones found more frequently across languages had a higher proportion of words that were correctly repeated, as clear in Figure 2.

```{r Fig2-xling_freq, ig.pos = "T!", fig.width=5, fig.height=5, fig.cap="NWR scores for individual target words as a function of the average frequency with which each phone is found across languages."}
aggregate(trnsc$correct*100,by=list(trnsc$target,trnsc$avg_fr),mean)->xling
colnames(xling)<-c("Target","Avg_fr","Word")
plot(xling$Word~xling$Avg_fr,lty=3,ylab="NWR scores (% words)",xlab="Average phone cross-ling frequency",type="n")
text(xling$Word~xling$Avg_fr,labels=xling$Target,cex=.5)
abline(lm(xling$Word~xling$Avg_fr))
```

@@

Next, we investigated patterns of deletion and substitution. Deletions were relatively rare, with only `r sum(del_tab[c(oralVs,nasalVs)])` vowels deleted, and `r sum(del_tab[c(simplexCs,complexCs)])` consonants.

@@

```{r bylength}
aggregate(trnsc$correct*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->by_length
aggregate(trnsc$pc*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->pc
aggregate(trnsc$nld*100,by=list(trnsc$id,trnsc$age.rounded,trnsc$type),mean)->nld
by_length=cbind(by_length,pc$x,nld$x)
colnames(by_length)<-c("ID","Age","Length","Word","Phoneme","NLD")

@@

Our final exploratory analysis assessed whether variation in scores was structured by factors that vary across individuals. As shown in Figure 2, there was a greater deal of variance in earlier than later ages, with significantly higher NWR scores for older children  (Spearman's rank correlation, given inequality of variance, rho (`r corage$statistic`) = `r corage$estimate`, p = `r corage$p.value`). In contrast, there was no clear association between NWR scores, on the one hand, and sex (t (`r tsex$statistic`) = `r tsex$parameter`, p = `r tsex$p.value`), birth order  (data missing for `r sum(is.na(demo$birthOrder[demo$included=="yes"]))` children, `r corbo$statistic`) = `r corbo$estimate`, p = `r corbo$p.value`) or maternal education  (data missing for `r sum(is.na(demo$matEd[demo$included=="yes"]))` children, `r corme$statistic`) = `r corme$estimate`, p = `r corme$p.value`)

### 2020-06-15 manuscript.tex
fixed little bugs, added freq ana and abstract

\abstract{
ADD
}

@@

\hypertarget{todo}{%
\subsection{TODO}\label{todo}}

\begin{itemize}
\tightlist
\item
  double check demo: exact age missing: 78, 82, 83, 115, 119; rough age missing 78, there's a mixture of BGMF in sex, first born is missing, round age
\item
  then add analysis of NWR as a function of frequency
\item
  draft discussion
\item
  draft abstract
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  look through for \textbf{mc}
\item
  double check my frequency entries in segments.xlsx using the search function in \url{https://phoible.org/parameters}; take a look at \url{http://phoible.github.io/conventions/} in case you see something about double articulation (I didn't find tp or lBj)
\item
  double check phonological implementation of stimuli in stimuli.xlsx
\item
  for both of the above, if you make any changes, they need to be re-saved as UTF16 txt file, and then converted (see remember above)
\item
  double check also stim\_seg\_freq -- why are some items still with NA freq?
\item
  consent, \enquote{correct}, \enquote{accuracy}, \enquote{performance}
\item
  help with table 1 (formatting \& making it appear, because it's broken by the phonological representations)
\end{itemize}

@@

NWR has been used to seek answers to a variety of theoretical questions, including what the links between phonology, working memory, and the lexicon are (Bowey, 2001). It is also frequently used as for applications, notably as a diagnostic for language delays and disorders (Estes, Evans, \& Else-Quest, 2007). Since non-words can be generated in any language, it has attracted the attention of researchers working in multilingual and linguistically diverse environments, particularly in Europe (Action, 2009; Meir, Walters, \& Armon-Lotem, 2016).

@@

@@

A total 55 children were tested, from 34 of families, in five hamlets. Some children could not be included for the following reasons: refused participation or failed to repeat items presented over headphones even after coaching (N=0), spoke too softly to allow offline coding (N=5). In addition, 2 teenagers were tested to put younger children at ease; their data is not included in analyses below. The remaining 40 children (14 girls) were aged 6.96 years (range 3.92-11.03 years). There were 32 children exposed only to Yélî in the home, 6 children who were also exposed to another language in the home, and , 2 for whom this information was missing. Maternal years of education averaged 8.24 years (range 6-12 years; 2 children had this information missing).\footnote{Education is often reported in even years because people typically complete two-year cycles.} In terms of birth order, 0 were first borns, 5 second, 4 third, 2 forth, 6 fifth, 5 sixth, and 1 did not have this information.

@@

We first checked whether accuracy varies between first and subsequent presentations of an item by averaging word-level accuracy at the participant level separately for first attempts and subsequent repetitions. We excluded 6 children who did not have data for one of these two types. As shown in Figure 1, participants' mean word-level accuracy became more heterogeneous in subsequent repetitions. Surprisingly, subsequent repetitions (M = 41, SD = 27)
were on average less accurate than first ones (M = 64, SD = 15), t(35) = 5.88, p = 0. Given the uncertainty in whether previous work used only the first or all repetitions, and since behavior degraded and became more heterogeneous in subsequent repetitions, the rest of the analyses focus on only the first repetitions.

Taking into account only the first attempts, we averaged attempts by each of the 24 children who had data for first attempts; their ages ranged from 3.92 to 10.20 (M = 6.50, SD = 1.50).

@@

We investigated patterns of deletion and substitution. Deletions were relatively rare, with only 29 vowels deleted, and 9 consonants.

@@

\begin{tabular}{lll}
\toprule
Word & \multicolumn{1}{c}{Phoneme} & \multicolumn{1}{c}{NLD}\\
\midrule
46 (22) & 58 (17) & 42 (17)\\
80 (23) & 94 (7) & 6 (7)\\
78 (20) & 94 (6) & 6 (6)\\
74 (30) & 93 (11) & 7 (11)\\
\bottomrule
\end{tabular}

@@

Our final exploratory analysis assessed whether variation in scores was structured by factors that vary across individuals. As shown in Figure 2, there was a greater deal of variance in earlier than later ages, with significantly higher NWR scores for older children (Spearman's rank correlation, given inequality of variance, rho (1,519.81) = 0.34, p = 0.10). In contrast, there was no clear association between NWR scores, on the one hand, and sex (t (0.18) = 21.34, p = 0.86), birth order (2,630.74) = -0.14, p = 0.50) or maternal education (2,091.92) = 0.09, p = 0.67)

@@

@@

>>>>

\abstract{
In nonword repetition (NWR) studies, participants are presented auditorily with an item that is phonologically legal but lexically meaningless in the local language. Accuracy is thought to reflect long-term phonological knowledge as well as online phonological working memory and flexible production patterns.In this study, we report on NWR results among children learning Yêly Dnyé, an isolate spoken in Rossel Island, PNG, with an unusually dense phonological inventory. This study contributes to three lines of research. First, we document that non-word items containing typologically rare sounds are repeated accurately less often that non-words containing more common sounds. Second, we document rather weak effects of item length, contributing to other research suggesting that length effects may be language-specific. Third, we do not find strong individual variation effects in this population, contrary to previous results documenting strong age-related effects. Together, these data provide a unique view of online phonological processing in a seldom-studied language, and contribute to both typological and language acquisition research.
}

@@

@@

\begin{itemize}
\tightlist
\item
  look through this manuscript for bits of text to be written by \textbf{mc}
\item
  help with table 1 (formatting \& making it appear, because it's broken by the phonological representations)
\item
  do you think Figure 2 should be done with phonetic symbols? If yes, could you try to implement that?
\item
  double check demo -- probably start with the xlsx version that has the children's first names, date of birth, and date of test. I looked again at my local files (still haven't checked the labs' ones) and I'm surprised I have maternal education for more children than I have birth order for\ldots{} If you correct this, to generate the csv I also round the age, assign age.rounded to reported age (in years) when DOB is missing, remove first names, DOB, DOTest and age.precise before pushing to github
\item
  double check my frequency entries in segments.xlsx using the search function in \url{https://phoible.org/parameters}; take a look at \url{http://phoible.github.io/conventions/} in case you see something about double articulation (I didn't find tp or lBj)
\item
  double check phonological implementation of stimuli in stimuli.xlsx
\item
  for both of the above, if you make any changes, they need to be re-saved as UTF16 txt file, and then converted (see remember above)
\item
  double check also stim\_seg\_freq -- why are some items still with NA freq?
\item
  consent, \enquote{correct}, \enquote{accuracy}, \enquote{performance}
\item
  draft discussion
\end{itemize}

@@

NWR has been used to seek answers to a variety of theoretical questions, including what the links between phonology, working memory, and the lexicon are (Bowey, 2001), and to assess the extent to which apparent phonological constraints found in the lexicon affect online production (Gallagher, 2014). It is also frequently used as for applications, notably as a diagnostic for language delays and disorders (Estes, Evans, \& Else-Quest, 2007). Since non-words can be generated in any language, it has attracted the attention of researchers working in multilingual and linguistically diverse environments, particularly in Europe (Action, 2009; Meir, Walters, \& Armon-Lotem, 2016).

@@

Additionally, we looked up each phoneme in each target word in the Phoible database (Moran \& McCloy, 2019). We entered the number of languages in which each phoneme was found, as well as what percentage of languages in the database that represented. We could not find a small number of phonemes in the database, and treat those as NA values that do not contribute to the target-level global average phone cross-linguistic frequency, although an alternative would be to attribute them a very small number. * \textbf{mc} do you agree with this decision or shall we do the alternative?*

@@

A total 55 children were tested, from 34 of families, in five hamlets. Some children could not be included for the following reasons: refused participation or failed to repeat items presented over headphones even after coaching (N=0), spoke too softly to allow offline coding (N=5). In addition, 2 teenagers were tested to put younger children at ease; their data is not included in analyses below. The remaining 40 children (14 girls) were aged 6.96 years (range 3.90-11 years). There were 32 children exposed only to Yélî in the home, 6 children who were also exposed to another language in the home, and , 2 for whom this information was missing. Maternal years of education averaged 8.24 years (range 6-12 years; 2 children had this information missing).\footnote{Education is often reported in even years because people typically complete two-year cycles.} In terms of birth order, 5 were first borns, 4 second, 2 third, 7 forth, 5 fifth, 1 sixth, and NA did not have this information.

@@

We first checked whether accuracy varies between first and subsequent presentations of an item by averaging word-level accuracy at the participant level separately for first attempts and subsequent repetitions. We excluded 3 children who did not have data for one of these two types. As shown in Figure 1, participants' mean word-level accuracy became more heterogeneous in subsequent repetitions. Surprisingly, subsequent repetitions (M = 39, SD = 26)
were on average less accurate than first ones (M = 65, SD = 15), t(38) = 6.62, p = 0. Given the uncertainty in whether previous work used only the first or all repetitions, and since behavior degraded and became more heterogeneous in subsequent repetitions, the rest of the analyses focus on only the first repetitions.

Taking into account only the first attempts, we averaged attempts by each of the 24 children who had data for first attempts; their ages ranged from 3.90 to 10.20 (M = 6.50, SD = 1.50).

@@

In this analysis, we were interested in variation in performance as a function of the average frequency with which sounds composing individual target words are found in languages over the world. To look at this, we fit a mixed logistic regression, in which the outcome variable was whether the non-word was correctly repeated or not. The fixed effects of interest was the average cross-linguistic phone frequency; we also included child age as a control fixed effect, and child ID \& target ID as random variables.

We could include 885 observations (\textbf{why -- shouldn't it be 1k??}), from 41 children producing in any given trial one of 41 potential target words. The analysis revealed a significant estimate for the average cross-linguistic frequency of phones in the target words (ß = 0.03, SE ß = 0.01, p = 0): Target words with phones found more frequently across languages had a higher proportion of words that were correctly repeated, as clear in Figure 2.

\begin{figure}
\centering
\includegraphics{manuscript_files/figure-latex/Fig2-xling_freq-1.pdf}
\caption{(\#fig:Fig2-xling\_freq)NWR scores for individual target words as a function of the average frequency with which each phone is found across languages.}
\end{figure}

Next, we investigated patterns of deletion and substitution. Deletions were relatively rare, with only 29 vowels deleted, and 9 consonants.

@@

\begin{tabular}{lll}
\toprule
Word & \multicolumn{1}{c}{Phoneme} & \multicolumn{1}{c}{NLD}\\
\midrule
47 (22) & 59 (17) & 41 (17)\\
79 (22) & 94 (7) & 6 (7)\\
78 (19) & 94 (6) & 6 (6)\\
75 (32) & 93 (11) & 7 (11)\\
\bottomrule
\end{tabular}

@@

Our final exploratory analysis assessed whether variation in scores was structured by factors that vary across individuals. As shown in Figure 2, there was a greater deal of variance in earlier than later ages, with significantly higher NWR scores for older children (Spearman's rank correlation, given inequality of variance, rho (1,548.04) = 0.33, p = 0.12). In contrast, there was no clear association between NWR scores, on the one hand, and sex (t (0.18) = 21.34, p = 0.86), birth order (data missing for 17 children, 2,678.47) = -0.16, p = 0.44) or maternal education (data missing for 2 children, 2,091.92) = 0.09, p = 0.67)

@@

\leavevmode\hypertarget{ref-gallagher2014identity}{}%
Gallagher, G. (2014). An identity bias in phonotactics: Evidence from Cochabamba Quechua. \emph{Laboratory Phonology}, \emph{5}(3), 337--378. \url{https://doi.org/10.1515/lp-2014-0012}

@@

\leavevmode\hypertarget{ref-phoible}{}%
Moran, S., \& McCloy, D. (Eds.). (2019). \emph{PHOIBLE 2.0}. Jena: Max Planck Institute for the Science of Human History. Retrieved from \url{https://phoible.org/}
